How functional traits influence plant growth and shade tolerance across the life-cycle

### 2018-04-15 ms/ms-supp.Rnw
Minor modifications for submission

\input{ms-common-defs}

@@

\hypersetup{
     linkcolor=pnasbluetext
}

\begin{document}

@@

% Figure command -- also adds figure to table of contents
\newcommand{\myfigure}[4]{
    \begin{figure*}[!hb]
            \centering
             \includegraphics[width=0.9\textwidth]{#1}
            \caption{\textbf{#2} #3}
            \label{#4}
    \end{figure*}
}

@@

Values of $p(\nu, E)$ are calculated across a range of values of $E$, and then an expression of the form in Eq. \ref{eq:photosynthesis} is fitted to extract the parameters
$\alpha_{{\rm p1}}$ and $\alpha_{{\rm p2}}$, such that these become functions of $\nu$.

@@

The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004},
\begin{equation}
k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.
\end{equation}
This relationship is normalised around $\phi_0$, the global mean of $\phi$. This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.

@@

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_core.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{7cm}llll',
            label="tab:params_core",
            caption="Core parameter of the \texttt{FF16} physiological model."),
      floating.environment="table*",
      sanitize.text.function=I,
      include.rownames=FALSE, 
      caption.placement = "top")
@

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_hyper.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{9cm}llll',
            label="tab:params_hyper",
            caption="Parameters for hyper-parameterisation of the \\texttt{FF16} physiological model."),
      sanitize.text.function=I,
      floating.environment="table*",
      include.rownames=FALSE, 
      caption.placement = "top")
@

@@

\myfigure{SI/lma_tradeoff.pdf}{Leaf turnover decreases with leaf-mass per unit leaf area.}{Data from \citep{Wright-2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to datsa from each site, with intensity of shading adjusted
according to strength of the relationship.}{fS-leaf}

@@

\myfigure{main/assumptions.pdf}{Assumptions of a functional balance model for plant construction.}{Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
Blue lines show standardised major axis lines fit to different species. The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.}{fig:assumptions}

@@

\myfigure{SI/growth_light_dia.pdf}{Effect of four traits on stem-diameter growth rate for different-sized plants.}{
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_dia}

\myfigure{SI/growth_light_area.pdf}{Effect of four traits on stem-area growth rate for different-sized plants.}{
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_area}

\myfigure{SI/growth_light_mass.pdf}{Effect of four traits on plant mass growth rate for different-sized plants.}{
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired mass and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}{fig:growth_light_mass}

>>>>

% abbreviations
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}
\newcommand{\sma}{\textsc{sma}}

\usepackage{graphicx}
\graphicspath{{../output/SI/}}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

% Please give the surname of the lead author for the running footer
\leadauthor{Falster}

@@

\hypersetup{
     linkcolor=pnasbluetext
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

@@

@@

The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004} (Fig. \ref{fig:leafturn}),
\begin{equation}
k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.
\end{equation}
This relationship is normalised around $\phi_0$, the global mean of $\phi$. This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.

@@

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Apr  5 13:08:59 2018
\begin{table*}[ht]
\centering
\caption{Core parameter of the 	exttt{FF16} physiological model.} 
\label{tab:params_core}
\begin{tabular}{p{7cm}llll}
  \hline
Description & Symbol & Unit & Code & Value \\ 
  \hline
\textbf{Plant construction} &  &  &  &  \\ 
  Crown-shape parameter & $\eta$ &  & \texttt{eta} & 12 \\ 
  Leaf mass per area & $\phi$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma} & 0.1978791 \\ 
  Wood density & $\rho$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho} & 608 \\ 
  Sapwood area per unit leaf area & $\theta$ &  & \texttt{theta} & 0.0002141786 \\ 
  Height of plant with leaf area of 1m$^2$ & $\alpha_{\rm l1}$ & m & \texttt{a\_l1} & 2.17 \\ 
  Exponent of relationship between height and leaf area & $\alpha_{\rm l2}$ &  & \texttt{a\_l2} & 0.5 \\ 
  Root mass per unit leaf area & $\alpha_{\rm r1}$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{a\_r1} & 0.07 \\ 
  Ratio of bark area to sapwood area & $\alpha_{\rm b1}$ &  & \texttt{a\_b1} & 0.17 \\ 
  \textbf{Production} &  &  &  &  \\ 
  Leaf photosynthesis per area & $\alpha_{\rm p1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-2}$ & \texttt{a\_p1} & 151.1778 \\ 
  Saturation of leaf photosynthesis per area & $\alpha_{\rm p2}$ &  & \texttt{a\_p2} & 0.2047162 \\ 
  Yield = fraction of carbon fixed converted into mass & $\alpha_{\rm y}$ &  & \texttt{a\_y} & 0.7 \\ 
  Biomass per mol carbon & $\alpha_{\rm bio}$ & $\mathrm{kg}\,\mathrm{mol}^{-1}$ & \texttt{a\_bio} & 0.0245 \\ 
  Leaf respiration per mass & $r_{\rm l}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_l} & 198.4545 \\ 
  Fine-root respiration per mass & $r_{\rm r}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_r} & 217 \\ 
  Sapwood respiration per mass & $r_{\rm s}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_s} & 6.598684 \\ 
  Bark respiration per mass & $r_{\rm b}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{r\_b} & 13.19737 \\ 
  Turnover rate for leaves & $k_{\rm l}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_l} & 0.1521952 \\ 
  Turnover rate for sapwood & $k_{\rm s}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_s} & 0.2 \\ 
  Turnover rate for bark & $k_{\rm b}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_b} & 0.2 \\ 
  Turnover rate for fine roots & $k_{\rm r}$ & $\mathrm{yr}^{-1}$ & \texttt{k\_r} & 1 \\ 
  \textbf{Fecundity} &  &  &  &  \\ 
  Seed mass & $\omega$ & kg & \texttt{omega} & 0.000038 \\ 
  Height at maturation & $H_{{\rm mat}}$ & m & \texttt{hmat} & 15 \\ 
  Maximum allocation to reproduction & $\alpha_{\rm f1}$ &  & \texttt{a\_f1} & 0.8 \\ 
  Parameter determining rate of change in $r(x, m_\textrm{l})$ around $H_{\rm mat}$ & $\alpha_{\rm f2}$ &  & \texttt{a\_f2} & 10 \\ 
   \hline
\end{tabular}
\end{table*}

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Apr  5 13:08:59 2018
\begin{table*}[ht]
\centering
\caption{Parameters for hyper-parameterisation of the \texttt{FF16} physiological model.} 
\label{tab:params_hyper}
\begin{tabular}{p{9cm}llll}
  \hline
Description & Symbol & Unit & Code & Value \\ 
  \hline
\textbf{Leaf turnover} &  &  &  &  \\ 
  Global average leaf mass per area & $\phi_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{lma\_0} & 0.1978791 \\ 
  Rate of leaf turnover at average leaf mass per unit leaf area, $\phi_0$ & $\beta_{\rm kl1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_kl1} & 0.4565855 \\ 
  Scaling exponent for $\phi$ in leaf turnover & $\beta_{\rm kl2}$ &  & \texttt{B\_kl2} & 1.71 \\ 
  \textbf{Sapwood turnover} &  &  &  &  \\ 
  Global average wood density & $\rho_0$ & $\mathrm{kg}\,\mathrm{m}^{-3}$ & \texttt{rho\_0} & 608 \\ 
  Rate of sapwood turnover at average wood density, $\rho_0$ & $\beta_{\rm ks1}$ & $\mathrm{yr}^{-1}$ & \texttt{B\_ks1} & 0.2 \\ 
  Scaling exponent for $\rho$ in sapwood turnover & $\beta_{\rm ks2}$ &  & \texttt{B\_ks2} & 1.25 \\ 
  \textbf{Photosynthesis} &  &  &  &  \\ 
  Leaf nitrogen per unit leaf area & $\nu$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea} & 0.00187 \\ 
  Global average nitrogen per unit leaf area & $\nu_0$ & $\mathrm{kg}\,\mathrm{m}^{-2}$ & \texttt{narea\_0} & 0.00187 \\ 
  Potential CO$_2$ photosynthesis at average leaf nitrogen, $\nu_0$ & $\beta_{\rm lf1}$ & $\mathrm{mol}\,\mathrm{d}^{-1}\,\mathrm{m}^{-2}$ & \texttt{B\_lf1} & 0.8273474 \\ 
  Curvature of light response curve & $\beta_{\rm lf2}$ &  & \texttt{B\_lf2} & 0.75 \\ 
  Quantum yield of leaf photosynthesis (CO$_2$ per unit photosynthetically active radiation) & $\beta_{\rm lf3}$ &  & \texttt{B\_lf3} & 0.04 \\ 
  Scaling exponent for leaf nitrogen in maximum leaf photosynthesis & $\beta_{\rm lf5}$ &  & \texttt{B\_lf5} & 0.5 \\ 
  \textbf{Respiration} &  &  &  &  \\ 
  CO$_2$ respiration per unit leaf nitrogen & $\beta_{\rm lf4}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{kg}^{-1}$ & \texttt{B\_lf4} & 15750 \\ 
  CO$_2$ respiration per unit sapwood volume & $\beta_{\rm rs1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rs1} & 4012 \\ 
  CO$_2$respiration per unit bark volume & $\beta_{\rm rb1}$ & $\mathrm{mol}\,\mathrm{yr}^{-1}\,\mathrm{m}^{-3}$ & \texttt{B\_rb1} & 8024 \\ 
   \hline
\end{tabular}
\end{table*}

\newpage
\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{lma_tradeoff.pdf}
\caption{\textbf{Leaf turnover decreases with leaf-mass per unit leaf area.}
Data from \citep{Wright-2004} for 678 species from 51 sites, each
point giving a species-average. Lines show standardised major axis lines
fitted to data from each site, with intensity of shading adjusted
according to strength of the relationship.}
\label{fig:leafturn}
\end{figure*}
\newpage

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{assumptions.pdf}
\caption{\textbf{Assumptions of a functional balance model for plant construction.}
Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
Blue lines show standardised major axis lines fit to different species. The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table 4 from main text.}
\label{fig:assumptions}
\end{figure*}

@@

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_dia.pdf}
\caption{\textbf{Effect of four traits on stem-diameter growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-diameter and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_dia}
\end{figure*}

@@

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_area.pdf}
\caption{\textbf{Effect of four traits on stem-area growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired stem-area and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_area}
\end{figure*}

@@

\begin{figure*}[!hb]
\centering
\includegraphics[width=0.9\textwidth]{growth_light_mass.pdf}
\caption{\textbf{Effect of four traits on plant mass growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table 3. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired mass and their growth rate estimated. Changes in trait-growth relationships are summarised in Table 5.}
\label{fig:growth_light_mass}
\end{figure*}

### 2018-04-15 ms/ms.tex
Minor modifications for submission

\input{ms-common-defs}

@@

% Enable cross referencing to Evidence document
\usepackage{xr}
\externaldocument{ms-supp}
\externaldocument{figure_conceptual}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{main/figure_conceptual.pdf}
\caption{Conceptual framework linking growth rate to plant size and traits.
\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant. In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$. Grey numbers indicate equation numbers referred to in the main text. The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}. Colours highlight where the same metric appears repeatedly in different equations. For a list of variable names see Table \ref{tab:definitions}.}
\label{fig:conceptual}
\end{figure}

\section*{Modeling the effects of traits on demography}

@@

% Update equation counter, following inclusion of figure 1
\setcounter{equation}{9}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.9\linewidth]{main/lcp_schematic.pdf}

@@

Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}. We assume the amount allocated to vegetative growth starts high amd then decreases once $H > H_{{\rm mat}}$  (Fig. \ref{fig:conceptual}b). The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction. Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.

@@

Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted accordingly whenever {\lma} is varied.

@@

\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
\centering
\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{main/growth_light_height.pdf}
\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated. The white regions indicate trait ranges that are typically observed in real systems. Figs. \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass. Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
\label{fig:growth_light_height}}
\end{SCfigure*}

@@

\begin{SCfigure*}[\sidecaptionrelwidth][b]
\centering
\includegraphics[width=11.4cm,keepaspectratio]{main/lcp_trait.pdf}
\caption{\textbf{Effect of three developmental traits on shade tolerance.}
Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero. Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m). The white regions indicate trait ranges that are typically observed in real systems.
\label{fig:wplcp}}
\end{SCfigure*}

@@

>>>>

\usepackage{graphicx}
\graphicspath{{../output/main/}}

% abbreviations
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}
\newcommand{\sma}{\textsc{sma}}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

% Please give the surname of the lead author for the running footer
\leadauthor{Falster}

@@

% Enable cross referencing to Evidence document
% \usepackage{xr}
% \externaldocument{ms-supp}
% \externaldocument{figure_conceptual}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figure_conceptual.pdf}
\caption{Conceptual framework linking growth rate to plant size and traits.
\textbf{a)} Shows how the distribution of mass in a typical plant varies with size.
\textbf{b)} Equations describing the rates of biomass growth, and growth in various dimensions of the plant. In the first line the symbol $\Sigma$ means ``sum'' across tissues, where $i = {\rm l,b,s,r}$. Grey numbers indicate equation numbers referred to in the main text. The insets show how the different metrics change intrinsically with plant height, when applying the ``functional-balance'' model in Table \ref{tab:allometry}. Colours highlight where the same metric appears repeatedly in different equations. For a list of variable names see Table \ref{tab:definitions}.}
\label{fig:conceptual}
\end{figure}

\section*{A framework for understanding the effects of traits on growth rate and shade tolerance}

@@

\refstepcounter{equation}\label{eq:dbdt_2}
\refstepcounter{equation}\label{eq:dMtdt}
\refstepcounter{equation}\label{eq:daldmt}
\refstepcounter{equation}\label{eq:daldt}
\refstepcounter{equation}\label{eq:dhdt}
\refstepcounter{equation}\label{eq:dast}
\refstepcounter{equation}\label{eq:dDdt}

% Update equation counter, following inclusion of figure 1
% \setcounter{equation}{9}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.9\linewidth]{lcp_schematic.pdf}

@@

Height at maturation: {\hmat} moderates an inevitable trade-off between mass invested reproduction and growing the plant, i.e.~ $\frac{{\rm d}M_{\rm a}}{{\rm d}B}$ in eqns \ref{eq:dhdt} and \ref{eq:dast}. We assume the amount allocated to vegetative growth starts high and then decreases once $H > H_{{\rm mat}}$  (Fig. \ref{fig:conceptual}b). The exact nature of the function is non-critical, what is important is that plants shift from a period of investing mainly in vegetative growth to investing mainly in reproduction. Here we let the fraction of mass allocated to reproduction vary with $H$ via the function $r_{\rm r1}  \left(1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)\right)^{-1}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.

@@

Leaf mass per unit leaf area: {\lma} (=$\phi$) influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link {\lma} to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship \citep{Wright-2004} (Fig. \ref{fig:leafturn}): $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\phi/\phi_0\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. The rate of leaf respiration per unit area is assumed independent of {\lma}\citep{Wright-2004}, as such the mass-based rate is adjusted whenever {\lma} is varied.

@@

\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
\centering
\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{growth_light_height.pdf}
\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated. The white regions indicate trait ranges that are typically observed in real systems. Figs. \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass. Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
\label{fig:growth_light_height}}
\end{SCfigure*}

@@

\begin{SCfigure*}[\sidecaptionrelwidth][b]
\centering
\includegraphics[width=11.4cm,keepaspectratio]{lcp_trait.pdf}
\caption{\textbf{Effect of three developmental traits on shade tolerance.}
Panels show effect of traits on level of canopy openness that causes biomass growth (eqn \ref{eq:dbdt}) to be zero. Different lines indicate relationship for plants with specified height, from short (light blue,  $H=0.5$m) to tall (dark line, $H=20$m). The white regions indicate trait ranges that are typically observed in real systems.
\label{fig:wplcp}}
\end{SCfigure*}

@@

% Defines references from supplemntary section
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

\refstepcounter{table}\label{tab:params_core}
\refstepcounter{table}\label{tab:params_hyper}
\refstepcounter{figure}\label{fig:leafturn}
\refstepcounter{figure}\label{fig:assumptions}
\refstepcounter{figure}\label{fig:growth_light_dia}
\refstepcounter{figure}\label{fig:growth_light_area}
\refstepcounter{figure}\label{fig:growth_light_mass}
